Creating the ``r params$package_name`` R package

### 2025-06-05 create-FSHybridPLS.Rmd
nipals first step done

## Documenting the package and building

>>>>

## PLS component computation

### get_pls_comp
Computes the first PLS component from a hybrid predictor and response vector using regularization.

- Inputs
  - **`W`**: A `predictor_hybrid` object with functional and scalar predictors.
  - **`y`**: A numeric response vector of length equal to `W$n_sample`.
  - **`L`**: A Cholesky decomposition of a positive definite regularization matrix.
- Output: 
  - `xi_hat`: The first PLS component as a `predictor_hybrid` object
  - `E`: The matrix `inv(L) %*% V_star %*% t(inv(L))`
  - `V_star`: Cross-covariance matrix between predictors and response
  - `eigen_val`: Leading eigenvalue of `E`
```{r}
#' Compute the First PLS Component from a Hybrid Predictor
#'
#' Computes the coefficients of the first Partial Least Squares (PLS) component based on 
#' a hybrid predictor object and a response vector, using a regularized generalized eigenvalue problem.
#'
#' @param W A \code{predictor_hybrid} object containing both functional and scalar predictors.
#' @param y A numeric response vector of length equal to the number of samples in \code{W}.
#' @param L cholesky decompisition of a regularization matrix (typically positive definite) used in the generalized eigenproblem.
#'
#' @return A list with the following elements:
#' \describe{
#'   \item{\code{xi_hat}}{The estimated first PLS component as a \code{predictor_hybrid} object.}
#'   \item{\code{E}}{The generalized eigenvalue problem matrix \code{E = inv(L) V* t(inv(L))}.}
#'   \item{\code{V_star}}{The cross-covariance matrix between predictors and response.}
#'   \item{\code{eigen_val}}{The leading eigenvalue of \code{E}.}
#' }
#'
#' @export
get_pls_comp <- function(W, y, L){

  V_star <- get_pre_corrcov(W, y)

  invL <- Matrix::chol2inv(L)
  A <- V_star %*% t(invL)       # A = V* × t(inv(L))
  E <- invL %*% A               # E = inv(L) × A

  eigen_result <- eigen(E)
  e <- eigen_result$vectors[, 1]
  if (is.complex(e)){
    print("stop")
    return("stop")
  }

  xi_star <- t(invL) %*% e      # xi_star solves t(L) xi = e
  xi_hat <- predictor_hybrid_from_coef(format = W, coef = xi_star)

  return(list(
    xi_hat = xi_hat,
    E = E,
    V_star = V_star,
    eigen_val = eigen_result$values[1]
  ))
}
```

```{r}
suppressPackageStartupMessages({
  library(fda)
  library(Matrix)
  library(ggplot2)
})

# --- Helper: Generate realistic smooth functional data ---
generate_fd_data <- function(n_sample, nbasis, freq = 1) {
  basis <- create.bspline.basis(c(0, 1), nbasis = nbasis, norder = 4)
  time_points <- seq(0, 1, length.out = nbasis)
  coefs <- outer(1:nbasis, 1:n_sample, function(i, j) {
    sin(freq * pi * time_points[i]) + rnorm(1, sd = 0.1)
  })
  fd(coef = coefs, basisobj = basis)
}

# --- Settings ---
set.seed(20250604)
n_sample <- 200
n_scalar <- 10
n_functional <- 2
nbasis_vals <- c(4,8,12,16,20)
lambda_vals <- c(0, 1e-8, 1e-6, 1e-4, 1e-2, 1, 10)

results <- data.frame()

# --- Experiment Loop ---
for (nbasis in nbasis_vals) {
  for (lambda in lambda_vals) {
    cat("Trying nbasis =", nbasis, ", lambda =", lambda, "\n")

    # Generate functional predictors
    fd1 <- generate_fd_data(n_sample, nbasis, freq = 1)
    fd2 <- generate_fd_data(n_sample, nbasis, freq = 2)
    Z <- matrix(rnorm(n_sample * n_scalar), n_sample, n_scalar)
    y <- rnorm(n_sample)

    W <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))
    lambda_vec <- rep(lambda, n_functional)

    # Try to compute constraint matrix L
    L <- get_constraint_matrix_hybrid(W, lambda_vec)
    constr_mat_chol <- t(chol(L))  # <-- do not change this

    # Time the PLS computation
    elapsed <- system.time({
      out <- tryCatch({
        pls <- get_pls_comp(W, y, constr_mat_chol)

        if (!is.list(pls) || (is.character(pls) && pls == "stop")) {
          data.frame(
            nbasis = nbasis,
            lambda = lambda,
            eigen_val = NA,
            is_real = FALSE,
            reason = "stop (complex eigenvector)",
            time_sec = NA_real_
          )
        } else {
          data.frame(
            nbasis = nbasis,
            lambda = lambda,
            eigen_val = Re(pls$eigen_val),
            is_real = TRUE,
            reason = "success",
            time_sec = NA_real_  # placeholder, will update below
          )
        }
      }, error = function(e) {
        data.frame(
          nbasis = nbasis,
          lambda = lambda,
          eigen_val = NA,
          is_real = FALSE,
          reason = "error in get_pls_comp",
          time_sec = NA_real_
        )
      })
    })

    out$time_sec <- elapsed["elapsed"]
    results <- rbind(results, out)
  }
}

# --- Print successful configurations ---
cat("\n✅ Successful runs with real eigenvalues:\n")
print(subset(results, is_real == TRUE))

# --- Fastest and slowest ---
cat("\n⏱️ Top 5 fastest settings:\n")
print(head(results[order(results$time_sec), ], 5))

cat("\n🐢 Top 5 slowest settings:\n")
print(tail(results[order(results$time_sec), ], 5))

# --- Visualize success/failure ---
ggplot(results, aes(x = nbasis, y = lambda, fill = reason)) +
  geom_tile(color = "white") +
  scale_y_log10() +
  scale_fill_manual(values = c(
    "success" = "green",
    "stop (complex eigenvector)" = "orange",
    "error in get_constraint_matrix_hybrid" = "red",
    "error in get_pls_comp" = "darkred"
  )) +
  labs(
    title = "get_pls_comp Outcomes: λ vs. # Basis Functions",
    x = "# of Basis Functions",
    y = "Smoothing Parameter λ (log scale)",
    fill = "Result Type"
  )

#

```{r eval=FALSE, include=FALSE}
nipals_pen_hybrid <- function(W, y, n_iter, lambda, verbose) {
  # 1. initialize the storage
  rho <- xi <- delta <- nu  <- sigma <- eta <- list()
  E <- V_star <- eigen_val <-list()
  fitted_value_W <- fitted_value_y <-list()
  resid_y <- W_now <- list() #data for iteration
  first_eigen_val <- mse_W <- mse_y <-rep(NA, n_iter)

  # 2. constraint matrix is the same at every iteration
  constr_mat      <- get_constraint_matrix_hybrid(W, lambda) 
  constr_mat_chol <- t(chol(constr_mat))

  W_now[[1]] <- W
  y_now <- y
  final_succesful_iteration <- 0
  for (l in 1:n_iter) {
    #cat(paste("#############################################", "\n"))
    if(verbose==T)cat(paste(l, "th component", "\n"))

    # STEP 1. calculate pls score
    pls_result_now <- get_pls_comp(W_now[[l]], y_now, constr_mat_chol, verbose)
    if(!is.list(pls_result_now)){break}

    xi[[l]] <- pls_result_now$xi # pls component. Section 3.2
    rho[[l]] <- hybrid_inner_prod(W_now[[l]], xi[[l]]) # pls scores : n_sample * 1 matrix

    # STEP 2. residulize
    # residualize predictor: hybrid (W) on scalar (rho) linear regression with n_sample observations
    delta[[l]] <- LSE_hybrid(W_now[[l]], rho[[l]]) #single obs of hybrid predictor

    fitted_value_W[[l]] <- fitted_value(delta[[l]], rho[[l]])
    W_now[[l+1]] <- subtr(W_now[[l]], fitted_value_W[[l]])
    # residualize resoponse
    nu[[l]] <- LSE_ptws(y_now, rho[[l]]) #scalar on scalar linear regression with n_sample observations
    fitted_value_y[[l]] <- nu[[l]] * rho[[l]]
    y_now <- y_now - fitted_value_y[[l]]
    resid_y[[l]] <- y_now

    #check mean norm
    #mse_W[l] <- mean_sum_sqrd(W_now[[l]])
    #mse_y[l] <- norm(y_now,"2")

    #cat("W:", mse_W[l], "\n")
    #cat("y:",  mse_y[l], "\n\n")

    # STEP 3.
    sigma[[l]] <- xi[[l]]
    if (l == 1) {
      eta[[l]] <- scalar_mul(sigma[[l]], nu[[l]])
    }else{ # if l > 1
      for (u in 1:(l - 1)){
        sigma[[l]] <- subtr( sigma[[l]], sigma[[u]], hybrid_inner_prod(delta[[u]], xi[[l]]) )
      } #end of u-loop
      eta[[l]] <- add(eta[[l-1]], sigma[[l]], nu[[l]])
    }#if statement

    # just for records.
    E[[l]] <- pls_result_now$E # for monitoring symetricity
    V_star[[l]] <- pls_result_now$V_star
    eigen_val[[l]] <- pls_result_now$eigen_val
    first_eigen_val[l] <- eigen_val[[l]][1]
    final_succesful_iteration <- l
  }# for loop: l
}
```
## Documenting the package and building
